Movie Recommendation System
a. Content-based recommendation systems
b. Collaborative filtering recommendation systems
c. (combine both) Hybrid recommendation

Dataset: MovieLens

#1.Importing libraries (packages: recommenderlab, ggplot2, data.table, reshape2')

```{r}
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
```

#2. Retrieving the Data
We will now retrieve our data from movies.csv into movie_data dataframe and ratings.csv into rating_data. We will use the str() function to display information about the movie_data dataframe.
```{r}
setwd("~/Documents/projects/netflix_R")
movie <- read.csv("movies.csv",stringsAsFactors = FALSE)
rating <- read.csv("ratings.csv")
str(movie)
```

```{r}
summary(movie)
```

```{r}
head(movie)
```
```{r}
summary(rating)
```
```{r}
head(rating)
```
#User Based Collaborative Filtering (UBCF) model
```{r}
rating
```

```{r}
ratingMat <- dcast(rating, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMat <- as.matrix(ratingMat[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMat <- as(ratingMat, "realRatingMatrix")
ratingMat
```

overview some of the important parameters that provide us various options for building recommendation systems for movies

```{r}
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
names(recommendation_model)
```

```{r}
lapply(recommendation_model, "[[", "description")
```

We will implement user-based collaborative filtering. (UBCF_realRatingMatrix)
```{r}
recommendation_model$UBCF_realRatingMatrix$parameters
```

```{r}
rec_mod = Recommender(ratingMat, method = "UBCF", param=list(method="Cosine",nn=10)) 
```

#movie predictions of the first user
```{r}
Top_5_pred = predict(rec_mod, ratingMat[1], n=5)
```
```{r}
Top_5_List = as(Top_5_pred, "list")
Top_5_List
```
# take a look at the movie names that correspond to these number
```{r}
movie
```


```{r}
library(dplyr)
Top_5_df=data.frame(Top_5_List)
colnames(Top_5_df)="movieId"
Top_5_df$movieId =as.numeric(Top_5_df$movieId)
names=left_join(Top_5_df, movie, by="movieId")
names
```

#3. Data pre-processing
one-hot encoding
```{r}
genre <- as.data.frame(movie$genres, stringsAsFactors = FALSE)
genre
```
```{r}
library(data.table)
genre2 <- as.data.frame(tstrsplit(genre[,1],'[|]', type.convert= TRUE), stringAsFactors = FALSE)
genre2
```
```{r}
unique(sort(genre2[,1]))
```

```{r}
colnames(genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre
str(genre_mat1)
```

```{r}
for (index in 1:nrow(genre2)) {
  for (col in 1:ncol(genre2)) {
    gen_col = which(genre_mat1[1,] == genre2[index,col]) 
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
genre_mat2

```
```{r}
SearchMatrix <- cbind(movie[,1:2], genre_mat2[])
head(SearchMatrix)
```


#4. Exploring Data
Collaborative Filtering provides recommendations based on the similar profiles of its users. One key advantage of collaborative filtering is that it is independent of the product knowledge. Rather, it relies on the users with a basic assumption that what the users liked in the past will also like in the future. For example, if a person A watches crime, sci-fi and thriller genres and B watches sci-fi, thriller and action genres then A will also like action and B will like crime genre.

```{r}
recommendation_model$UBCF_realRatingMatrix$parameters
```

```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ],
                               method = "cosine",
                               which = "users")
as.matrix(similarity_mat)

image(as.matrix(similarity_mat), main = "User's Similarities")
```
```{r}
movie_similarity <- similarity(ratingMatrix[, 1:4], method =
                                 "cosine", which = "items")
as.matrix(movie_similarity)

image(as.matrix(movie_similarity), main = "Movies similarity")
```
```{r}
rating_values <- as.vector(ratingMatrix@data)
unique(rating_values) # extracting unique ratings
```
```{r}
Table_of_Ratings <- table(rating_values) # creating a count of movie ratings
Table_of_Ratings
```

#5. Most Viewed Movies Visualizstion
 We will first count the number of views in a film and then organize them in a table that would group them in descending order.
```{r}
library(ggplot2)
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie,
                                         movie$movieId == table_views[index,1])$title)
}
table_views[1:6,]
```
```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'pink') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  ggtitle("Total Views of the Top Films")
```
#6. Heatmap of Movie Ratings
```{r}
image(ratingMatrix[1:20, 1:25], axes = FALSE, main = "Heatmap of the first 25 rows and 25 columns")
```
#7. Performing Data Preparation
a. Selecting useful data.
b. Normalizing data.
c. Binarizing the data.
For finding useful data in our dataset, we have set the threshold for the minimum number of users who have rated a film as 50. This is also same for minimum number of views that are per film. This way, we have filtered a list of watched films from least-watched ones.

```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                             colCounts(ratingMatrix) > 50]
movie_ratings
```

```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                     colCounts(movie_ratings) > minimum_users],
main = "Heatmap of the top users and movies")
```

```{r}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("steelblue"), col=I("red")) +
  ggtitle("Distribution of the average rating per user")
```
Data Normalization
```{r}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)

image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                          colCounts(normalized_ratings) > minimum_users],
main = "Normalized Ratings of the Top Users")
```
Data Binarization
```{r}
binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.95)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.95)
#movies_watched <- binarize(movie_ratings, minRating = 1)

good_rated_films <- binarize(movie_ratings, minRating = 3)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
colCounts(movie_ratings) > binary_minimum_users],
main = "Heatmap of the top users and movies")
```
#8. Collaboratice filtering system
This type of collaborative filtering finds similarity in the items based on the people’s ratings of them. The algorithm first builds a similar-items table of the customers who have purchased them into a combination of similar items. This is then fed into the recommendation system.

The similarity between single products and related products can be determined with the following algorithm –

For each Item i1 present in the product catalog, purchased by customer C.
And, for each item i2 also purchased by the customer C.
Create record that the customer purchased items i1 and i2.
Calculate the similarity between i1 and i2.
We will build this filtering system by splitting the dataset into 80% training set and 20% test set.
```{r}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```
```{r}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters
```

```{r}
recommen_model <- Recommender(data = training_data,
                          method = "IBCF",
                          parameter = list(k = 30))
recommen_model
class(recommen_model)
```
Using the getModel() function, we will retrieve the recommen_model. We will then find the class and dimensions of our similarity matrix that is contained within model_info. Finally, we will generate a heatmap, that will contain the top 20 items and visualize the similarity shared between them.
```{r}
model_info <- getModel(recommen_model)
class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
   main = "Heatmap of the first rows and columns")
```
In the next step of ML project, we will carry out the sum of rows and columns with the similarity of the objects above 0. We will visualize the sum of columns through a distribution as follows –
```{r}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)

sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red"))+ ggtitle("Distribution of the column count")
```

```{r}
top_recommendations <- 10 # the number of items to recommend to each user
predicted_recommendations <- predict(object = recommen_model,
                          newdata = testing_data,
                          n = top_recommendations)
predicted_recommendations
```

```{r}
user1 <- predicted_recommendations@items[[1]] # recommendation for the first user
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie,
                                         movie$movieId == movies_user1[index])$title)
}
movies_user2
```
```{r}
recommendation_matrix <- sapply(predicted_recommendations@items,
                      function(x){ as.integer(colnames(movie_ratings)[x]) }) # matrix with the recommendations for each user

recommendation_matrix[,1:4]
```

